RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-10-14 doc/paper.tex
Main RILS-ROLS algorithm pseudocode finished.
Some fixing in pseudocode.
Update: OLS section

 \subsection{Ordinary least square method}
   Ordinary least square method (OLS) is a linear regression technique. It is based on applying the least-square method to minimize the square residual (error) sum  between actual and predicted values (given by the model). More precisely, given data $({X}, y)$, the task is to determine linear mapping $\tilde{y} = k \textbf{x} + b$, that is coefficients (line slope) $k = (k_1, \ldots, k_d)$ and $b$ (intercept), so that $ \sum_{i} |\tilde{y}_i - y_i|^2 $ is minimized. This sum is also know as the sum of squared error (SSE). There are many methods to minimize SSE. One of the analytical approaches is the calculus-oriented -- it takes into account   partial derivation w.r.t. $k$ of the cost function $ \sum_{i} |\tilde{y}_i - y_i|^2 $, getting

   $$  \frac{\partial}{\partial k} \sum_{i} |\tilde{y}_i - y_i|^2 = \frac{\partial}{\partial k} \sum ( kx_i+b  - y_i)^2 =  \sum -2x_i(kx_i + b - y_i)  =  \sum -2x_i (\tilde{y}_i - y_i ) . $$

   Taking into account   partial derivation w.r.t.  $b$, we have 
    $$  \frac{\partial}{\partial b} \sum_{i} (\tilde{y}_i - y_i)^2  = \frac{\partial}{\partial b} \sum ( kx_i+b  - y_i)^2 = \sum -2  (\tilde{y}_i - y_i). $$

    Thus, we get the system of equation: 
    \begin{align*}
    	  &\sum -2x_i (\tilde{y}_i - y_i )  = 0 \\
    	  &\sum -2  (\tilde{y}_i - y_i ) = 0.
    \end{align*}
     The above system can be transformed to 
     \begin{align*}
     	 & \sum y_i \cdot x_i   = b\sum x_i \cdot x_i  +  k \cdot \sum x_i \\
     	 & \sum y_i  = k \sum x_i + b \cdot n. 
     \end{align*}
     By solving this system of equations, one can easily obtain vector of coefficients $k=(k_1, ..., k_d)$ and intercept $b$. \fxnote{TODO: Aca -- ovde je pisalo da je to sistem sa dve jednacine i dve nepoznate k i b -- medjutim, k je vektor. Ako mozes dopisi jos nesto, tipa koliko je podataka potrebno i slicno -- mozda veremenska slozenost.) }
   \subsection{RILS-ROLS method}
   Now we will explain the proposed RILS-ROLS method in detail. The overall method scheme is given in Algorithm~\ref{alg:rilsrols}.

>>>>

 \subsection{Ordinary least square method}
   Ordinary least square method (OLS) is a linear regression technique. It is based on applying the least-square method to minimize the square residual (error) sum  between actual and predicted values (given by the model). More precisely, given data $({X}=(X_1, \ldots, X_d)^T, y)$, the task is to determine linear mapping $\tilde{y} = k \textbf{x} + b$, that is coefficients (line slope) $k = (k_1, \ldots, k_d)$ and $b$ (intercept), so that $ \sum_{i} |\tilde{y}_i - y_i|^2 $ is minimized. This sum is also know as the sum of squared error (SSE). There are many methods to minimize SSE. One of the analytical approaches is the calculus-oriented -- it takes into account   partial derivation w.r.t. $k$ of the cost function $ \sum_{i} |\tilde{y}_i - y_i|^2 $, getting

   $$  \frac{\partial}{\partial k} \sum_{i} |\tilde{y}_i - y_i|^2 = \frac{\partial}{\partial k} \sum ( kX_i+b  - y_i)^2 =  \sum 2X_i(kX_i + b - y_i)  =  \sum -2X_i (y_i - \tilde{y}_i) . $$

   Taking into account   partial derivation w.r.t.  $b$, we have 
    $$  \frac{\partial}{\partial b} \sum_{i} (\tilde{y}_i - y_i)^2  = \frac{\partial}{\partial b} \sum ( kX_i+b  - y_i)^2 = \sum -2  (y_i - \tilde{y}_i). $$

    Thus, we get the system of equation: 
    \begin{align*}
    	 % &\sum -2X_i (\tilde{y}_i - y_i )  = 0 \\
    	  & -2X_1 (y_1 - \tilde{y}_1)  = 0 \\
    	  & \vdots \\
    	  & -2X_d (y_d - \tilde{y}_d)  = 0 \\
    	  &\sum -2  (y_i - \tilde{y}_i) = 0.
    \end{align*}
     %The above system can be transformed to 
     %\begin{align*}
     %	 & \sum y_i \cdot x_i   = b\sum x_i \cdot x_i  +  k \cdot \sum x_i \\
     %	 & \sum y_i  = k \sum x_i + b \cdot d. 
     %  \end{align*}
     By solving this system of $(d+1)$ equations, one can obtain vector of coefficients $k=(k_1, ..., k_d)$ and intercept $b$. %For example, in order to calculate coefficient $k_i$, the following system of equations is 
     \fxnote{TODO: Aca -- ovde je pisalo da je to sistem sa dve jednacine i dve nepoznate k i b -- medjutim, k je vektor. Ako mozes dopisi jos nesto, tipa koliko je podataka potrebno i slicno -- mozda veremenska slozenost.) }%https://towardsdatascience.com/understanding-the-ols-method-for-simple-linear-regression-e0a4e8f692cc
   \subsection{RILS-ROLS method}
   Now we will explain the proposed RILS-ROLS method in detail. The overall method scheme is given in Algorithm~\ref{alg:rilsrols}.
